SubGenome Analysis

### 2017-07-26 R/SubGenome Analysis.Rmd
Added data files and separating functions
Starting to separate functions into individual files, preparing to set this up as a single pipeline.

---
title: "SubGenome Analysis"
output: html_document
---

@@

## R Markdown
Some preprocessing was needed to generate the input files for this project.  We begin with SynMap output based on the Schnable 2011 paper (see https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3 for details).

This resulted in the file (31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks).

We parse this file to a flat format using a perl script (parseKsKnFile.pl)

@@

> perl parseKsKnFile.pl ../Data/31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum\ vs\ maize.tab

```{r parseSynMap, include=FALSE}
cmd <- paste("perl", "parseKsKnFile.pl", "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks")
system(cmd)
```

@@

## Import data and remove incomplete rows
syntelogs.import <- read_delim("../Data/sorghum vs maize.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]

@@

```{r writeSubgenomesToFile, echo=FALSE}
## Now recover which genes are in these sets
# subset(syntelogs.orthologs, count>=12 & median<=ks_cutoff & org_chr1=="a31607_1" & org_chr2=="b34889_1")
# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=ks_cutoff,]
subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
write.table(subgenome, "../Output/recoveredSubGenomes.tab", sep="\t")
```

>>>>

---
title: "SubGenome Analysis"
output: html_document
params:
  fileParser: "parseKsKnFile.pl"
  inputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
---

@@

## Maize v1 vs. Sorghum v1 (Based on Schnable 2011 paper)
Some preprocessing was needed to generate the input files for this project.  We begin with SynMap output based on the Schnable 2011 paper (see https://www.evernote.com/shard/s303/nl/52987123/1ce74f11-98a0-4854-b363-3172e79eeae3 for details).

This resulted in the file (6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks) comparing sorghum v1 to maize v1.

We parse this file to a flat format using a perl script (parseKsKnFile.pl) and manually correct the header lines.

@@

> perl parseKsKnFile.pl ../Data/6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks > ../Data/sorghum_v1_vs_maize_v1.tab

```{r parseSynMap, include=FALSE}
cmd <- paste("perl", params$fileParser, params$inputFile)
system(cmd)
```

@@

## Import data and remove incomplete rows
syntelogs.import <- read_delim("../Data/sorghum_v1_vs_maize_v1.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]

@@

```{r writeSubgenomesToFile, echo=FALSE}
## Now recover which genes are in these sets
# subset(syntelogs.orthologs, count>=12 & median<=ks_cutoff & org_chr1=="a31607_1" & org_chr2=="b34889_1")
# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=ks_cutoff,]
subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
write.table(subgenome, "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab", sep="\t")
```

We also want to know how we did compared to the paper itself, to see if this method is faithfully reproducing their results.
```{r compareResultsToPaper, echo=FALSE}
library(readxl)
library(splitstackshape)
#sd01 <- read_excel("~/git/SubGenomes/Data/sd01.xls")
#sd01$Maize1[grepl("^GRMZM", sd01$Maize1) | grepl("^AC", sd01$Maize1)]
```

## Maize v4+rejected set vs. Sorghum v3.1 (updated to use up-to-date assemblies and gene models)

```{r parseSynMap2, include=FALSE}
cmd <- paste("perl", "parseKsKnFile.pl", "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks")
system(cmd)
```

```{r importSynMapData2, echo=FALSE}
## Import data and remove incomplete rows
syntelogs.import <- read_delim("../Data/sorghum_v3.1_vs_maize_v4+rejected.tab", "\t", escape_double = FALSE, trim_ws = TRUE)
syntelogs.complete <- syntelogs.import[complete.cases(syntelogs.import),]

```{r pickKSCutoff2, echo=FALSE}
## Pick a ks value which separates alpha duplication from beta duplication (between the first two peaks)
log10_ks_cutoff <- -0.575
qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
ks_cutoff <- 10^log10_ks_cutoff
#ks_cutoff <- 0.4234 #include most of the first 2 peaks
```

Find subgenomes

```{r findSubgenomes2, echo=FALSE}
## Find which sets of chromosomes with syntelogs should be in group 1 or group 2, where group 1 has larger syntenic blocks
# First group rows with same syntenic block and chromosome, then summarize each block's ks values with median/mean/count
syntelogs.aggregate <- ddply(syntelogs.complete, ~block+org_chr1+org_chr2, summarise, median=median(ks), mean=mean(ks), count=length(ks))

Write to file

```{r writeSubgenomesToFile2, echo=FALSE}
## Now recover which genes are in these sets
# subset(syntelogs.orthologs, count>=12 & median<=ks_cutoff & org_chr1=="a31607_1" & org_chr2=="b34889_1")
# orthologs.blocks_groups <- syntelogs.orthologs[syntelogs.orthologs$count>=12 & syntelogs.orthologs$median<=ks_cutoff,]
subgenome.groups <- rbind(merge(syntelogs.orthologs, orthologs.sub1, all.x = F), merge(syntelogs.orthologs, orthologs.sub2, all.x = F))
subgenome <- merge(syntelogs.complete, subgenome.groups, all.x = F)
write.table(subgenome, "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab", sep="\t")
```
